---
title: "Exploring the Dependence of Fuel Economy on Type of Transmission"
author: "Sergey Cherkasov"
date: "18 Sep 2015"
output: pdf_document
---

#Synopsis
In this paper we are going to explore the relationship between a set of variables and miles per gallon (MPG) (outcome). We use mtcars dataset, provided by The Comprehensive R Archive Network and can code book can be found [here](http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html) - http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html.

The client is particularly interested in the following two questions:

* Is an automatic or manual transmission better for MPG  
* Quantify the MPG difference between automatic and manual transmissions

Original Rmd file could be found at the folloing GitHub [repository](https://github.com/PestoVerde/RegModels_PeerAssessment1) - https://github.com/PestoVerde/RegModels_PeerAssessment1.

#Exploratory analysis 

First let us investigate the data. It discribes `r dim(mtcars)[1]` cars with `r dim(mtcars)[2]` variables. These are `r names(mtcars)`. The parameters we are interesting for are mpg and am. Mean for automatic transmission is `r mean(mtcars$mpg[mtcars$am==0])` mpg whereas mean for manual transmission equals `r mean(mtcars$mpg[mtcars$am==1])`. Ranges are `r range(mtcars$mpg[mtcars$am==0])` and `r range(mtcars$mpg[mtcars$am==1])`, they do not even intersect.(see also boxplot in Appendix 1).

The dependence looks significant, and test comfirm that, with p-value:
```{r echo=FALSE}
auto <- mtcars$mpg[mtcars$am==0]
manual <- mtcars$mpg[mtcars$am==1]
test <- t.test(auto, manual)
test$p.value
```


##Is an automatic or manual transmission better for MPG? 



##Quantify the MPG difference between automatic and manual transmissions 



#Conclusion 


#Appendices

##Appendix 1
```{r echo=FALSE}
boxplot(mpg ~ as.factor(am), data=mtcars,
        main = "Dependance MPG on Type of Transmission",
        xlab = "Transmission", ylab = "MPG")
```
